A Theory for Data Migration of Generated Evolving Information Systems

### 2022-10-30 2022 Migration/articleMigration.tex
WIP

   Concepts are names that group atoms of the same type.
   For example, an engineer might choose to classify \atom{Peter} and \atom{Melissa} as \concept{Person},
   and \atom{074238991} as a \concept{TelephoneNumber}.
   In this example, \concept{Person} and \concept{TelephoneNumber} are concepts.
   We will use variables $A$, $B$, $C$, $D$ to represent concepts.
   The expression $a\ \inst\ A$ means that atom $a$ is an \emph{instance} of concept $A$.
   All concepts are taken from an infinite set $\Concepts$.
   $\Concepts$ and $\Atoms$ are disjoint sets.

@@

   Triples serve to represent data.
   A dataset $\dataset$ is a set of triples.
   A triple is an element of $\Triple{\Atoms}{\Rels}{\Atoms}$.
   This makes our theory valid for any kind of database that triples can represent,
   such as SQL databases, object-oriented databases, graph databases, triple stores, and other no-SQL databases
   Each triple relates two atoms to a relation.
   A triple $\triple{\text{\atom{Peter}}}{\id{phone}}{\text{\atom{074238991}}}$ might mean in practice that
   the ``thing'' that \atom{Peter} refers to has \atom{074238991} as a telephone number.
   This ``meaning from practice'' has no consequences in the formal world.
   We leave it entirely up to a user to attach a practical meaning to a triple.
   In this paper, the informal meaning is interesting only for the sake of illustration.
   To save writing in the sequel, we will write $a\ r\ b$ to denote that $\triple{a}{r}{b}\in\dataset$
   in those cases where $\dataset$ is either clear from the context or irrelevant for the argument.
   A relation $r$ can serve as a container of pairs,
   as defined by the funciton $\id{pop}_r:\Dataset\rightarrow\Pair{\Atoms}{\Atoms}$.
   It defines a set of pairs, which we call the population of $r$:
\begin{equation}
   \pop{r}{\dataset}\ =\ \{ \pair{a}{b}\mid\ \triple{a}{r}{b}\in\dataset\}
\end{equation}

@@

   This section introduces the notion of information system as a quintuple $\la\dataset_{\infsys},\schema_{\infsys},\inst_{\infsys},\roles_{\infsys},\maintain_{\infsys}\ra$.
   As before, we will omit the suffixes if $\infsys$ is obvious (i.e. clear from the context or irrelevant for the argument).
\begin{definition}[information system]
\label{def:information system}
\item An information system $\infsys$ is a tuple $\la\dataset,\schema,\inst,\roles,\maintain\ra$, in which
\begin{itemize}
   \item dataset $\dataset$ is defined as in section~\ref{sct:Datasets};
   \item schema $\schema$ is defined as in sections~\ref{sct:Schemas};
   \item $\inst : \Atoms\times\Concepts$ is the instance relation between atoms from $\dataset$ and concepts from $\schema$;
   \item $\roles$ is a set of roles;
   \item $\maintain : \roles\times\rules$ is the maintainance relation between roles and rules from $\schema$; and
   \item requirements~\ref{eqn:inst} thru~\ref{eqn:specialization} are satisfied.
\end{itemize}
\end{definition}
   Let us discuss these items one by one.

>>>>

   Concepts are names that group atoms of the same type.
   For example, an engineer might choose to classify \atom{Peter} and \atom{Melissa} as \concept{Person},
   and \atom{074238991} as a \concept{TelephoneNumber}.
   In this example, \concept{Person} and \concept{TelephoneNumber} are concepts.
   We will use variables $A$, $B$, $C$, $D$ to represent concepts.
   The expression $a\ \inst\ C$ means that atom $a$ is an \emph{instance} of concept $C$.
   The relation $\inst:\Pair{\Atoms}{\Concepts}$ is the instance relation between atoms and concepts.
   All concepts are taken from an infinite set $\Concepts$.
   $\Concepts$ and $\Atoms$ are disjoint sets.

@@

   Triples serve to represent data.
   A triple is an element of $\Triple{\Atoms}{\Rels}{\Atoms}$.
   This makes our theory valid for any kind of database that triples can represent,
   such as SQL databases, object-oriented databases, graph databases, triple stores, and other no-SQL databases
   Each triple relates two atoms to a relation.
   A triple $\triple{\text{\atom{Peter}}}{\id{phone}}{\text{\atom{074238991}}}$ might mean in practice that
   the ``thing'' that \atom{Peter} refers to has \atom{074238991} as a telephone number.
   This ``meaning from practice'' has no consequences in the formal world.
   We leave it entirely up to a user to attach a practical meaning to a triple.
   In this paper, the informal meaning is interesting only for the sake of illustration.
   To save writing in the sequel, we will write $a\ r\ b$ to denote that $\triple{a}{r}{b}\in\dataset$
   in those cases where $\dataset$ is either clear from the context or irrelevant for the argument.

   A dataset $\dataset$ is a tuple $\pair{\triples_{\dataset}}{\inst_{\dataset}}$,
   in which $\triples_{\dataset}$ is a finite set of triples
   and $\inst_{\dataset}$ is the instance relation.
   We will simply write $\triples$ or $\inst$ if the dataset is clear from the context or irrelevant for the argument.

   A relation $r$ can serve as a container of pairs,
   as defined by the function $\id{pop}_r:\Dataset\rightarrow\Pair{\Atoms}{\Atoms}$.
   It defines a set of pairs, which we call the population of $r$:
\begin{equation}
   \pop{r}{\dataset}\ =\ \{ \pair{a}{b}\mid\ \triple{a}{r}{b}\in\triples\}
\end{equation}

@@

   This section introduces the notion of information system as a quadrtuple $\la\dataset_{\infsys},\schema_{\infsys},\roles_{\infsys},\maintain_{\infsys}\ra$.
   As before, we will omit the suffixes if $\infsys$ is obvious (i.e. clear from the context or irrelevant for the argument).
\begin{definition}[information system]
\label{def:information system}
\item An information system $\infsys$ is a tuple $\la\dataset,\schema,\roles,\maintain\ra$, in which
\begin{itemize}
   \item dataset $\dataset$ is defined as in section~\ref{sct:Datasets};
   \item schema $\schema$ is defined as in sections~\ref{sct:Schemas};
   \item $\roles$ is a set of roles;
   \item $\maintain : \roles\times\rules$ is the maintainance relation between roles and rules from $\schema$; and
   \item requirements~\ref{eqn:inst} thru~\ref{eqn:specialization} are satisfied.
\end{itemize}
\end{definition}
   Let us discuss these items one by one.
